---
title: "General Social Survey"
author: "Kelly Douglas"
date: "Year 2019"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
Gss <- read.spss("GSS2018.sav")
Gss1 <- as.tibble(Gss)
```

```{r include=FALSE,echo=FALSE}
#names(Gss1)
mydata <- select(Gss1,VOTE16,RACE,REGION,EDUC,BIGBANG,EVOLVED,DEGREE,POSTLIFE,AGE,OWNGUN,CAPPUN,COURTS,POLVIEWS,VOTE12,GOD)
mydata$AGE <- as.numeric(mydata$AGE)
mydata$EDUC <- as.numeric(mydata$EDUC)
mydata2 <-filter(mydata,RACE!="IAP" ) %>% droplevels()
mydata3 <- filter(mydata2,BIGBANG=="True" | BIGBANG=="False",EVOLVED=="True"|EVOLVED=="False",POSTLIFE=="YES"|POSTLIFE=="NO",OWNGUN=="YES"|OWNGUN=="NO",COURTS=="TOO HARSH"|COURTS=="NOT HARSH ENOUGH"|COURTS=="ABOUT RIGHT") %>% droplevels()
mydata4 <- filter(mydata3,CAPPUN=="FAVOR"|CAPPUN=="OPPOSE",VOTE12=="Voted"|VOTE12=="Did not vote",VOTE16=="Voted"|VOTE16=="Did not vote") %>% droplevels()
levels(mydata4$VOTE12)[1] <- "voted12"
levels(mydata4$VOTE12)[2] <- "no in 12"

```



# Introduction Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

This data is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 15 variables. It has further removed unwanted factor levels in much of the data. Examine that code to see what it is doing. Finally address this question.  Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?
$H_0$: There is no difference in a persons opinion on the death penalty and if they own a gun. 

$H_A$: If a person owns a gun it does affect if they believe in the death penalty.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
For this problem I will be using a chisq test and a barchart to show you the results for the question of do people who own a gun believe in the death penalty. 
##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
barchartGC(~CAPPUN+OWNGUN,data=mydata4)
barchartGC(~CAPPUN+OWNGUN,data=mydata4,type="percent")
```

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=mydata4)
rowPerc(table1)
colPerc(table1)
```
Based on the information in the table it shows that more people that own guns favor the death penalty. 
### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
Based on the chisq test on this question the pvalue is 0.01651 and so I fail to reject the null hypothesis.
# Introduction Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Now answer Is belief in the bigbang theory (BIGBANG) independent of belief in the afterlife (POSTLIFE)
$H_0$: There is no difference in a persons opinion on the big bang theory and if they believe in the post life. 

$H_A$: If a person believes in the bigbang theory they also believe in the postlife.
## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
For this problem I will be using a chisq test and a barchart to show you the results for the question of is belief of the bigbang theory independent of belief in the afterlife. 
##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results
```{r}
barchartGC(~BIGBANG+POSTLIFE,data=mydata4)
barchartGC(~BIGBANG+POSTLIFE,data=mydata4,type="percent")
```

Based on the bar chart above it shows the big bang theory is independent to believing in the post life.
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results
```{r}
table1 <- xtabs(~BIGBANG + POSTLIFE, data=mydata4)
rowPerc(table1)
colPerc(table1)
```
Based on the table above the results are close in yes it is independent of each other. 
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
Basedc off the chisq test the p value is 0.08. So I would reject the null hypothesis. 
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

# Introduction Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Finally determine if a persons opinion on death penalty (CAPPUN) is independent of their race (RACE).
$H_0$: There is no difference in a persons opinion on the death penalty and their race. 

$H_A$: The persons race determines if they believe in the death penalty.
## Methods
I will be using the bar chart and the chisq test to see if there is a difference in this information. 
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

```{r}
barchartGC(~CAPPUN+RACE,data=mydata4)
barchartGC(~CAPPUN+RACE,data=mydata4,type="percent")
```
Based on the bar chart above there is a very big difference. It shows that more white people are in favor of the death penalty.
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

#### Numerical Descriptive Results
```{r}
table1 <- xtabs(~RACE + CAPPUN, data=mydata4)
rowPerc(table1)
colPerc(table1)
```
Based on the table above it shows that a persons race does affect their decision on their belief on the death penalty. 
<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

### Inferential Results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
Based on the chisq test the p value is 0.001282 so I would reject the null hypotesis that there is no connection between the race and if they believe in the death penalty. 
<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion
I have used bar charts and chisq tests to determine the p-values in these questions and decided if I was going to reject the null hypothesis or fail to reject the null hypothesis. I have responded to each of these at the end of each question. 
<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

